\section{Lecture 10}
On isomorphism theorems. The first isomorphism in action.
\begin{proposition}
  A subgroup $H\leq G$ is normal if and only if $H$ is the kernel of some homomorphism
  $\phi: G\to G'$.
\end{proposition}
\begin{proof}
  Note that every kernel is a normal subgroup by the first isomorphism theorem.
  Conversely, if $N\trianglelefteq G$, we can define the quotient map $\phi:G\to
  G/N, \text{ } g\mapsto gN.$ 
  This is a group homomorphism with kernel $N$ since $\phi(g) = 1_{G/N} \iff gN = N \iff g\in N.$
  %all $g\in G$ s.t. $\phi(g)=N$, which is all $g\in N.$ In other words, $\ker \phi=N.$
\end{proof}

\begin{remark}
    For a group $G,$ it is essentially equivalent to list normal subgroups as it is to list group homomorphisms out of $G.$
\end{remark}
% Then, taking the quotient by a normal subgroup $N$ is essentially equivalent to applying a
% homomorphism from $G$ to some other gorup whose kernel is $N$.
\begin{definition}
  Let $(H,*),(K,\cdot)$ be groups. We define their \emph{direct product} as the group
  with underlying set $H\times K = \left\{ (h,k) : h\in H, k\in K \right\}$ with operation
  of point-wise multiplication
  $(h,k)(h',k')=(h*h',k\cdot k').$
  \label{directProduct}
\end{definition}


\begin{proposition}
    Let $H$ and $K$ be groups, and let $G=H \times K$. The subset $H \times\left\{1_K\right\}$ is a normal subgroup of $G$ isomorphic to $H$, and one has $G / H \times\left\{1_K\right\} \cong K$. Similarly, $\left\{1_H\right\} \times K$ is a normal subgroup of $G$ isomorphic to $K$, and one has $G /\left(\left\{1_H\right\} \times K\right) \cong H$.
\end{proposition}

\begin{proof}
    Consider the map
    $$\phi :H\times \{1_K\} \to H$$ defined by $\phi ((h,1))=h.$ This is clearly an isomorphism.
    Now consider the canonical projection map
    $$\pi: H\times K \to K$$ defined by $\pi ((h,k))=k.$
    Then $\ker \pi =\{(h,1_K):h\in H\}=H\times 1_K.$ Since kernels are normal subgroups, we have shown that $H\times /{1_K}$ is a normal subgroup of $G.$ We have $\operatorname{Im} \pi =K$ trivially. Therefore, by the FIT, we have $H\times K / H\times \{1_K\} \cong K,$ or equivalently, $G/H\times {1_K} \cong K.$ 
    Similarly, we have $\{1_H\}\times K \cong K,$ is a normal subgroup of $G,$ and $G /\left(\left\{1_H\right\} \times K\right) \cong H.$
\end{proof}

\begin{example}
  We have previously seen (Example \ref{D12}) that there exists a homomorphism from $D_{12}$ to $S_3$. The kernel of 
  this homomorphism is the subgroup $H$ generated by the rotation by $180^o$, and the
  image is all of $S_3$. By FIT, we have $D_{12}/H \cong S_3$.
\end{example}
\begin{example}
  Let $\CC^{\times}$ be the non-zero complex numbers as a group under multiplication. The
  function $f:\RR\to\CC, \text{ } x\mapsto e^{2\pi i x}$ is a group homomorphism. The image is the
  unit circle in $\CC$, denoted $S^1$, and the kernel is $\ZZ$. Then,
  \[\RR/\ZZ \cong S^1.\]
  This can be extended to show that 
  \[\RR^n/\ZZ^n \cong \bigoplus_{i=1}^n S^1.\]
\end{example}

\begin{example}
  Consider the homomorphism from $\CC^{\times}$ to $\RR_{>0}$ viewed as groups under multiplication given by $z\mapsto |z|$. The image is $\RR_{>0}$ and the kernel is
  $S^1$. Hence by FIT,
  \[\CC^{\times} / S^1 \cong \RR_{>0}\]
\end{example}
\begin{example}
  How many distinct homomorphisms $\phi:G\to G'$ are there from $G=C_4=\langle g \mid g^4=1\rangle$ to $ G'= C_{10}=\langle
  g' \mid (g')^{10}=1\rangle$?  

 We will begin by listing our options for possible kernels and images, and then classify our homomorphisms by the choices of kernel.
First note that since $G$ is cyclic, it is abelian and therefore all of its subgroups are normal and can be kernels of the map.

The group $G$ has three subgroups: the trivial subgroup $\{e\},$ the subgroup $\langle g^2 \rangle =\{e,g^2\},$ and the group itself, $G.$

Our options for image are subgroups of $G'.$ We have: $\{e\}$, $\langle (g')^5 \rangle =\{e,(g')^5\},$ and the group itself, $G'.$

Now consider a map $\phi$ with $\ker \phi =\{e\}$. By FIT, we must have $G/\{e\} \cong G \cong \operatorname{Im} \phi,$ but $G'$ has no subgroup of order 4, so there is no such map.\\ Now consider $\phi$ with $\ker \phi =\{e,g^2\}.$ Again, by FIT, we must have $C_4/\{1,g^2\}=\{\{1,g^2\},\{g,g^3\}\} \cong \operatorname{Im}.$ We can take $\operatorname{Im} \phi =\{e,(g')^5\}$.\\ Finally, consider $\ker \phi =G.$ We must have $G/G \cong \{e\} \cong \Img\phi.$ We can take $\Img\phi =\{e\},$ and this is in fact the trivial homomorphism sending every element to the identity in $G'.$ \\Therefore, there are only two possible homomorphisms.
\end{example}
